---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ShinyFriendlyCaptcha

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ShinyFriendlyCaptcha)](https://CRAN.R-project.org/package=ShinyFriendlyCaptcha)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R badge](https://img.shields.io/badge/Build%20with-♥%20and%20R-blue)](https://github.com/mhanf/ShinyFriendlyCaptcha)
[![R-CMD-check](https://github.com/mhanf/ShinyFriendlyCaptcha/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mhanf/ShinyFriendlyCaptcha/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ShinyFriendlyCaptcha is to provide [Friendly Captcha](https://friendlycaptcha.com/) to Shiny apps.

## Features

Friendly Captcha is an European alternative to Google Recaptcha. It allow to protect websites against spam and bots in a privacy-embedded design.

Friendly Captcha features :

* Cryptographic bot protection
* No labeling tasks for users
* No tracking and cookies
* Fully accessible
* Guaranteed availability with SLA
* GDPR compliance agreements

ShinyFriendlyCaptcha additional features :

* Theme optimized for Bootstrap 5 delivered by the [bslib](https://rstudio.github.io/bslib/) package.
* Compatible with server validation of inputs based on [shinyvalidate](https://rstudio.github.io/shinyvalidate/) package.

## Installation

You can install the development version of ShinyFriendlyCaptcha from [GitHub](https://github.com/) with:

``` {r installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("mhanf/ShinyFriendlyCaptcha")
```
## Example 

In order to use the ShinyFriendlyCaptcha library, you’ll need to subscribe a friendly Captcha plan to obtain valid `SITEKEY` and `SECRET` strings. A free plan is proposed to developers for non-commercial use (protection of 1 website with up to 1,000 requests/month). For more details go to their [website](https://friendlycaptcha.com/).

ShinyFriendlyCaptcha exports two main functions: `sfc_output()` and `sfc_server()`. A vignette is available [here]() to help in the proper setup of these functions. A minimalistic Shiny app example using these functions can be created as follows:
 
``` {r example, eval=FALSE}
# libraries
library(shiny)
library(ShinyFriendlyCaptcha)
library(bslib)
library(shinyjs)
library(shinyvalidate)
# simple card function
card_template <- function(id, title, body){
  # card
  div(
    class = "card", 
    id = id,
    # card header
    div(class = "card-header text-center", title),
    # card body
    div(class = "card-body", body)
  )
}
# UI
ui <- fluidPage(
  theme = bs_theme(version=5, bootswatch = "flatly"),
  useShinyjs(),
  br(),
  fluidRow(
    column(width = 3, class = "mx-auto",
           # hidden final message
           hidden(
             h4(
               id ="final-msg", 
               class = "text-primary text-center", "Thank you !"
               )
           ),
           # contact form
           card_template(
             id = "form-contact",
             title = "Contact form",
             body = tagList(
               textInput(
                 inputId = "name",
                 label = "Name",
                 width = "100%"
               ),
               textInput(
                 inputId = "surname",
                 label = "Surname",
                 width = "100%"
               ),
               textInput(
                 inputId = "mail",
                 label = "Mail",
                 width = "100%"
               ),
               textAreaInput(
                  inputId = "msg",
                 label = "Message",
                 width = "100%",
                 height = "150px"
               ),
               sfc_output(
                 id = "test",
                 sitekey = Sys.getenv("captcha_sitekey"),
                 lang = "en",
                 dark_mode = FALSE,
                 eu_endpoint = FALSE,
                 theme_bs5 = TRUE
               ),
               br(),
               actionButton(
                 inputId = "ok",
                 label= "Submit",
                 width = "100%",
                 class="bg-primary"
               )
             )
           )
    )
  )
)
# Server
server <- function(input, output) {
  # shinyvalidate
  iv <- InputValidator$new()
  iv$add_rule("name", sv_required())
  iv$add_rule("surname", sv_required())
  iv$add_rule("mail", sv_required())
  iv$add_rule("mail", sv_email())
  iv$add_rule("msg", sv_required())
  iv$add_rule(
    "test-captchaId", 
    sv_equal(TRUE,message_fmt ="Captcha validation required")
    )
  # captcha response
  captcha_result <- sfc_server(
    id = "test",
    secret = Sys.getenv("captcha_secret"),
    sitekey = Sys.getenv("captcha_sitekey"),
    eu_endpoint = FALSE
  )
  # action on button click
  observeEvent(input$ok,{
    iv$enable()
    req(iv$is_valid())
    req(captcha_result()$success)
    hide(id = "form-contact")
    show(id="final-msg")
  })
}
# Run the application
shinyApp(ui = ui, server = server)
```

## Contributions

The developer and maintainer of ShinyFriendlyCaptcha is [mhanf](https://github.com/mhanf). External contributions are welcome. Please keep in mind that I am not a professional R developer but an enthusiastic R data scientist who plays with shiny and javascript as a pretext to learn new stuffs. Unfortunately, so is my code. Please note that the ShinyFriendlyCaptcha project is released with a [Contributor Code of Conduct](https://mhanf.github.io/ShinyFriendlyCaptcha/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## License

The ShinyFriendlyCaptcha package as a whole is licensed under the [MIT](https://opensource.org/licenses/mit-license.php) license.



